---
title: "Recruitment"
subtitle: Finding Future Leaders
---

# Recruiting our students {#recruitment}
Many of our students are recruited _before_ they graduate by working for an employer on a fixed-term contract, either part-time, full-time and as summer interns or for a year long placement with industrial experience, aka sandwich degree program.

![Are these graduates the future of your organisation? If you want them to be, you need to engage early and often with students *before* they graduate. This will increase your chances of recruiting them. The picture shows University of Manchester graduates celebrating their graduation outside the [Samuel Alexander](https://en.wikipedia.org/wiki/Samuel_Alexander) building with photobombing by [Gavin Brown](https://profgavinbrown.github.io/)](employing-your-future.png)


## About our students {#students}

We are one of the biggest departments of Computer Science at a [Russell Group](https://www.russellgroup.ac.uk/) University with a lots of students. As of 2026 we have *over 1000 undergraduate students* studying one of [two main Bachelors degrees](https://www.cs.manchester.ac.uk/study/undergraduate/courses/):

1. BSc (Hons) Computer Science 
1. BSc (Hons) Computer Science with Mathematics

Both degree programes are available *with Industrial Experience* (IE), a year in industry. Alongside our undergraduate community we currently have:

* Over 240 postgraduate taught students, doing Masters degrees
* Over 240 PhD students doing research supervised by academics in the Department

## Year long placements {#ie}

Around a third of our undergraduate students choose to do a year in industry. For the academic year starting in  2025, we have 77 students starting a year long industrial experience (IE) placement as part of their degree.

```{r bigie-fig, echo = FALSE, results='hide', message=FALSE, warning=FALSE, fig.align = "center", out.width = "100%", fig.cap = "Number of students starting an Industrial Experience (IE) placement year in 2024 as part of their undergraduate study at the University of Manchester. The data is plotted by Department: `CS`: Department of Computer Science [cs.manchester.ac.uk](https://www.cs.manchester.ac.uk/), `MACE`: Department of Mechanical, Aerospace and Civil Engineering (MACE) [mace.manchester.ac.uk](https://www.mace.manchester.ac.uk/), `EEE`: Department of Electrical and Electronic Engineering [eee.manchester.ac.uk](https://www.eee.manchester.ac.uk/), `ChemEng`: Department of Chemical Engineering [ce.manchester.ac.uk](https://www.ce.manchester.ac.uk/), `Chemistry`: Department of Chemistry [chemistry.manchester.ac.uk](https://www.chemistry.manchester.ac.uk/), `Maths`: Department of Mathematics [maths.manchester.ac.uk](https://www.maths.manchester.ac.uk), `Physics`: Department of Physics & Astronomy [physics.manchester.ac.uk](https://www.physics.manchester.ac.uk), `Materials`: Department of Materials [materials.manchester.ac.uk](https://www.materials.manchester.ac.uk/), `EES`: Department of Earth and Environmental Sciences [ees.manchester.ac.uk](https://www.ees.manchester.ac.uk/). Data from PowerBI as of 2025, note that summer internships are shown in this bar chart, only year long placements."}

library(tidyverse)
library(ggplot2)
library(forcats)

ieuom <- readr::read_csv("https://raw.githubusercontent.com/dullhunk/cdyf/refs/heads/master/data/ieuom.csv")

ggplot(ieuom, aes(y = forcats::fct_rev(forcats::fct_infreq(Department)))) +
  geom_bar() +
  labs(y = "Department at the University of Manchester", x = "Number of undergraduate students starting IE placements in 2024")
```

<!--
Still working out how to use ggplot with uomie.csv
See https://github.com/dullhunk/cdyf/issues/1187
Current solution is to make the data more like mpg with ieuom.csv
-->


We are working hard to improve the employability of students because while having a Computer Science is necessary for some jobs, it is not sufficient. Over the last decade we have been successful in *more than doubling* the number of our students going on year long placements in industry to around 100 per year, see figure \@ref(fig:ie-fig). This is a win-win-win situation for:

1. **Students**: benefit from a broader education, and develop softer social and non-cognitive skills that can be challenging to teach and learn in a purely academic environment.
2. **Employers**: placements are a cost-effective way for employers to recruit (and retain) graduate talent
3. **Universities**: produce better graduates with broader and deeper skills, who earn more and get better jobs. [@graduateoutcomes]



```{r ie-fig, echo = FALSE, results='hide', message=FALSE, warning=FALSE, fig.align = "center", out.width = "100%", fig.cap = "Number of undergraduate Computer Science students completing a year in industry as part of their degree at the University of Manchester. Since 2008 over 900 students have completed the program. I have been leading the program since 2012 and managed to ~double the number of students doing placements per year. As you can see in the histogram, the COVID-19 pandemic started having an effect on students starting placements in 2020."}

library(tidyverse)
library(ggplot2)

evenmoarplacements <- readr::read_csv("https://raw.githubusercontent.com/dullhunk/cdyf/master/data/onie.csv")


ggplot(data = evenmoarplacements) + geom_bar(mapping = aes(x = year, y = students), stat = "identity") + labs(y = "Number of students", x = "Start of Industrial Experience (IE) Year")
```

Previously our students have secured year long placements at a wide range of employers including Accenture, Agilent Technologies, Amazon, AND Digital, Apadmi, Arggo, ARM, Autodesk, AVL Powertrain, BAML, the BBC, Biorelate, BJSS, Bloomberg, BMW Mini, Bsquare Controls, BT, Cantarus, Celtra, CERN, Codethink, d3t, Elysian Systems, Feral Interactive, Fidelity, FiveAI, HMRC, IBM, Imagination Technologies, Intel, ISA Software, JP Morgan, Keysight Technologies, KPMG, Matillion, McAfee, Mentor Graphics, Monoprix, Morgan Stanley, NCC Group, Nokia, Nomura, Novacoast, Ocado, PA Consulting, PwC, Schlumberger, ServiceNow, Siemens, Soda Software, SteamaCo, The Hut Group, The Start Up Factory, Uber, Visa and Vodafone.



## Summer internships {#interns}
Many of our students do sumer internships betweens 1st and 2nd, or 2nd and 3rd year of study. Unlike year-long placements, these are not a formal part of their undergraduate degree program, but we encourage our students to take skilled work during summer.

## Part-time work 
Some of our students do part-time work during term-time, for example through Imago, the student software company [imago.cs.manchester.ac.uk](https://imago.cs.manchester.ac.uk/) as well as other employers.

## Wednesday Waggle {#waggle}

During term time, we highlight events and vacancies for Computer Science students in a weekly jobs newsletter called the *Wednesday Waggle*. 

This newsletter goes out to 1000+ Bachelors and Masters Computer Science students in Manchester each week. If you have vacancies or events you would like our students to know about, you can advertise them at [careerconnect.manchester.ac.uk](https://careerconnect.manchester.ac.uk/). 

Alternatively, point us to where they are advertised online, see [advertising vacancies to students](https://www.cdyf.me/finding#studentjobs). It can help students find your needle in the jobs haystack.

## Student societies

The University of Manchester Students Union (UMSU) is the largest in the UK with over 500 student societies. Many of these societies have a computational, technical and interdisciplinary interests including (in no particular order):

* Startups [accelerateme.co](https://accelerateme.co)
* University Computer Science Manchester [unicsmcr.com](https://unicsmcr.com/) 
* Manchester University Data Science & Artificial Intilligence [mudsai.com](https://www.mudsai.com)
* IBM Z society [IBMZ.uom](https://linktr.ee/ibmz.uom)
* Maths Society [MathSoc](https://manchesterstudentsunion.com/activities/view/mathsoc)
* Manchester Intelligence Society (Cybersecurity) [MIS](https://manchesterstudentsunion.com/activities/view/manchester-intelligence)
* Robosoc [uom-robosoc.com](https://uom-robosoc.com)
* Hyperloop Manchester [hyperloopmanchester.co.uk](https://www.hyperloopmanchester.co.uk)
* Makerspace Manchester [makerspacemanchester.org](https://www.makerspacemanchester.org)
* Manchester Stinger Motorports [formulastudentuom.com](https://formulastudentuom.com)
* Google Developer Group (GDG) Manchester [GDGManchester](https://linktr.ee/gdgmanchester)

A good way to collaborate with and recruit students is to sponsor and support their events by contacting relevant societies directly. This includes hackathons, talks, panels, workshops, insights and more. 

If you are recruiting computer scientists and software engineers as a summer interns, placement students or as graduates please get in touch. We typically have around 250 undergraduate students graduating annually, alongside a smaller number of Masters and PhD students. As of 2026, the [entry tariff](https://www.ucas.com/ucas/tariff-calculator) of our students (A* A* A* including mathematics) is comparable to other leading Computer Science (CS) departments in [Russell Group](https://en.wikipedia.org/wiki/Russell_Group) universities.

## Beyond Computer Science

<!-- was an entry tariff table here-->

If you are looking to recruit or collaborate with science and engineering students from other disciplines like [Physics](https://www.physics.manchester.ac.uk/), [Maths](https://www.maths.manchester.ac.uk/), [Chemistry](https://www.chemistry.manchester.ac.uk/), Mechnical, Aerospace & Civil Engineering ([MACE](https://www.mace.manchester.ac.uk/)), [Materials Science](https://www.materials.manchester.ac.uk/) and Electrical & Electronic Engineering ([EEE](https://www.eee.manchester.ac.uk/)) you should talk to:  

* academic staff in those departments and/or  
* the central careers service of the University via [employers.manchester.ac.uk](http://www.employers.manchester.ac.uk/)
* business engagement and knowledge Echange team [manchester.ac.uk/collaborate](https://www.manchester.ac.uk/collaborate/)

